\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-03 europa-basics.tex
updates

\comment{fix \newline and numbering problem above}

@@

\paragraph \texttt{Scheduling}: In the Resource Constrained Project
Scheduling Problem (RCPSP), a project consisting of a set of activities
must be scheduled in a way that satisfies minimum and/or maximum
temporal separation constraints. The activity schedule must also respect
fixed limits on the availability of resources required to perform each
activity. In addition to satisfying temporal and resource constraints,
it is common for the user to want to minimize makespan so that the
entire project is finished as early as possible. Fig. \ref{fig:rcpsp-1}
shows an example of a a solution provided by \eu for an RCPSP instance
with 10 activities, 5 resources and 30 temporal constraints.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small RCPSP problem}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

\paragraph \texttt{Planning}: In the Shopping Agent Problem
\cite{russelnorvig} an agent needs to purchase a set of products (milk,
drill, etc) that are available at specific locations (supermarket,
hardware store, etc), the agent is subject to temporal (must complete
tasks by specific deadlines) and resource (fuel, carrying capacity, etc)
constraints. The agent must figure out what actions need to be performed
to find and acquire the required items, as well as when to perform each
of those actions.

Below is a solution produced by \eu for a simple problem instance where
the shopping agent needs to buy Bananas, Milk and a Drill.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-Shopping0.jpg}
\caption{\small Shopping Agent problem}
\label{fig:shopping-1}
\vskip+0.1cm
\end{figure}

As we can see, a planning problem (what actions must be performed to
achieve a goal) may embed a scheduling problem (in what order should
those actions be performed), and both planning and scheduling embed
Constraint Satisfaction problems since temporal, resource and many other
kinds of constraints must be respected in most real life problems.

These relationships between Planning, Scheduling and Constraint
Satisfaction have been examined in detail in the literature (TODOL ref
Frank et al on Planning vs. Scheduling), and they lead to \eu's use of
constraint reasoning as its innermost building block. Below we examine
the main ideas from Constraint Satisfaction Programming, and from
Constraint Based Planning and Scheduling that constitute \eu's
theoretical underpinnings, then we dive into how \eu implements those
ideas.

@@

Constraint Satisfaction Programming, also know as Constraint Programming (CP) is a discipline that provides a generic framework for representing, solving and making logical inference on constraints. A complete treatment of this discipline is given in [MAR98] and [APT03] and very concise introductions are provided in [BAR99] and [LUS01].

	A constraint programming problem consists of a set of variables V={x1,..,xN}, where each variable takes values from a domain d1,..,dN. Domains are most often discrete and finite, but there are numerous CP implementations associated with continuous domains. Given a defined conjunctive set of constraints on the variables: $C=\{c1(x1,..,xN), ..., cK(x1,..,xN)\}$, the objective is to find one or more value assignments to V where all constraints are satisfied.

To solve a problem, CP uses logical inference to perform 2 operations:

\begin{enumerate}
	\item Bounds propagation: To infer upper and lower variable bounds. For example, from the constraints x1 + x2 <= 2, xi>=0, we can infer [0,2] bounds for x1 and x2

	\item Domain reduction: To infer a valid set of values for a variable. For example, for discrete variables, the constraints allDifferent(x1,x2,x3) and xi>=0, xi <=4. if x1 =1 and x2=3 we can infer that the valid domain for x3 is reduced to {2,4}

@@

CP is normally implemented as part of a programming language and constraints are normally represented as objects [PUG95]. Any constraint can be introduced and, as long as the bounds propagation and domain reduction protocols specified by the host CP system are enforced, it will be indistinguishable from any other ÒprimitiveÓ CP constraint, such as <= or >=.

Solving a CSP is NP-Hard in theory, but often very efficient in practice using a number of algorithms and techniques that are well understood.

TODO: talk about arc consistency algorithms, computational complexity and solvers.

@@

The most common planning formulations use a propositional representation, where the state of the world is represented by a set of propositions (statements that can be true or false), and operators change the truth values of these propositions (TODO ref?). Although these formulations are powerful and have allowed researchers to develop numerous contributions in automated planning, there are many classes of problem domains that are difficult it is difficult to represent using them.

In particular, it is hard to represent time, resources, mutual exclusion and concurrency using propositions (TODO: ref?).  It is straightforward to represent and reason about all of those elements using a CSP representation, as a result there has been some work on doing automated planning while taking advantage of CSP  (TODO: ref).  Traditionally, the entire planning problem is translated into a CSP and then solved using traditional CSP methods, this leads to a formulation where action choices and relationships are represented as variables and/or constraints. However, this has at least two major drawbacks:
\begin{enumerate}
	\item Given that action choices and relationships (rules) are expresed through variables, the domain descriptions that result from this approach are not intuitive and therefore difficult to understand and debug
	\item If the structure of a planning model (actions, conditions, effects, dependencies at the action level, etc) is not explicitly maintained by the CSP planner, the search algorithms are deprived of critical information to make better decisions. If that structure is maintained somehow (for instance, by internally marking variables that represent action choice and relationships between actions), it is still hard to write search algorithms as any planning-specific insight has to be translated into the CSP representation of variables and constraints  (TODO: example?)
\end{enumerate}

Constraint-Based Attribute and Interval planning (CAIP) (TODO:ref) is intended to close that gap: on the one hand, it takes advantage of a CSP representation, but it also uses attributes and intervals to maintain an explicit representation of the elements relevant to planning, so that it is easier to write algorithms that search for an reason abut plans. The main elements of CAIP are:

\begin{enumerate}
	\item Intervals:
	\item Attributes:
	\item Domain Constraints an Configuration rules:
\end{enumerate}

TODO: Explain how planning problem instances and their solutions are represented.

>>>>

\comment{fix newline and numbering problem above}

@@

\paragraph \texttt{Scheduling}: In the Resource Constrained Project
Scheduling Problem (RCPSP) \comment{citations needed}, a project
consisting of a set of activities must be scheduled in a way that
satisfies minimum and/or maximum temporal separation constraints. The
activity schedule must also respect fixed limits on the availability of
resources required to perform each activity. In addition to satisfying
temporal and resource constraints, it is common for the user to want to
minimize makespan so that the entire project is finished as early as
possible. Fig. \ref{fig:rcpsp-1} shows an example of a a solution
provided by \eu for an RCPSP instance with 10 activities, 5 resources
and 30 temporal constraints.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an RCPSP \comment{citation} problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

\paragraph \texttt{Planning}: In the Shopping Agent Problem
\cite{russelnorvig} an agent needs to purchase a set of products (milk,
drill, etc) that are available at specific locations (supermarket,
hardware store, etc), the agent is subject to temporal (must complete
tasks by specific deadlines) and resource (fuel, carrying capacity, etc)
constraints. The agent needs to figure out what actions need to be
performed to find and acquire the required items, as well as when to
perform each of those actions. Fig. \ref{fig:shopping-1} shows a
solution produced by \eu for such a problem instance.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-Shopping0.jpg}
\caption{\small A \eu solution to a shopping agent problem domain where
  the agent needs to buy Bananas, Milk and a Drill.}
\label{fig:shopping-1}
\vskip+0.1cm
\end{figure}

As the examples above show, a planning problem (actions to achieve a
goal) may embed a scheduling problem (what resources are necessary to
achieve stated goals) and both planning and scheduling could embed a
Constraint Satisfaction problem. % since temporal, resource and  other
% kinds of constraints must be respected in most real life problems.
The relationships between Planning, Scheduling and Constraint
Satisfaction have been examined \cite{smith00} and have lead to use of
constraint reasoning in \eu as its innermost building block. % Below we
% examine the main ideas from Constraint Satisfaction Programming, and
% from Constraint Based Planning and Scheduling that constitute
% \texttt{EUROPA}$_2$'s theoretical underpinnings.

@@

Constraint Satisfaction Programming (also known as Constraint
Programming (CP)) is a discipline that provides a generic framework for
representing, solving and making logical inference on constraints. A
complete treatment of this discipline is given in [MAR98] and [APT03]
and very concise introductions are provided in [BAR99] and
[LUS01].\comment{fix citations}

A constraint programming problem consists of a set of variables $V=
{x_1,..,x_N}$, where each variable takes values from a domain
$d_1,..,d_N$; in this chapter we will deal only with discrete and
finite domains. Given a defined conjunctive set of constraints on the
variables: $C=\{c_1(x_1,..,x_N), ..., c_K(x_1,..,x_N)\}$, the objective
is to find one or more value assignments to $V$ where all constraints
are satisfied.

To solve a problem, CP techniques use logical inference to perform two
important operations:

\begin{enumerate}

\item \textbf{Bounds propagation}: To infer upper and lower variable
  bounds. For example, from the constraints $x_1 + x_2 \leq 2$ \&
  $x_i \geq 0$, we can infer $[0,2]$ bounds for $x_1$ and $x_2$

\item \textbf{Domain reduction}: To infer a valid set of values for a variable.
  For example, for  constraints allDifferent($x_1,x_2,x_3$), $x_i \geq
  0$ \& $x_i \leq 4$, if $x_1 = 1$ and $x_2 = 3$ we can infer that the
  valid domain for $x_3$ is reduced  to $\{2,4\}$

@@

CP is normally implemented as part of a programming language and
constraints are normally represented as objects [PUG95]
\comment{citation}. Any constraint can be introduced and, as long as the
bounds propagation and domain reduction protocols specified by the host
CP system are enforced, it will be indistinguishable from any other
``primitive'' CP constraint, such as $\leq$ or $\geq$. In theory,
solving a CSP problem is NP-Hard \cite{ghallab04}, but often very
efficient in practice using a number of algorithms and techniques that
are well understood.

\comment{TODO: talk about arc consistency algorithms, computational
  complexity and solvers.}

@@

The most common planning formulations use a propositional
representation, where the state of the world is represented by a set
of propositions (statements that can be true or false), and operators
change the truth values of these propositions \cite{gen87}. Although
these formulations are powerful and have allowed researchers to
develop numerous contributions in automated planning, there are many
classes of problem domains that are difficult to represent using this
formalism. In particular, it is hard to represent time, resources,
mutual exclusion and concurrency using propositions \comment{citation
  needed}. While CSP representations have traditionally had an edge in
formulating and representing planning problems, 
% It is straightforward to represent and reason about all of
% those elements using a CSP representation, as a result there has been
% some work on doing automated planning while taking advantage of CSP
% (TODO: ref). Traditionally, the entire planning problem is translated
% into a CSP and then solved using traditional CSP methods, this
leading to a formulation where action choices and relationships are
represented as variables and/or constraints, there are at least two
major drawbacks:

\begin{enumerate}

\item Given that action choices and relationships (rules) are expressed
  through variables, the domain descriptions that result from this
  approach are not intuitive and therefore difficult to understand and
  debug

\item If the structure of a planning model (actions, conditions,
  effects, dependencies at the action level) is not explicitly
  maintained by the CSP planner, the search algorithms are deprived of
  critical information to make better decisions. If that structure is
  maintained (for instance, by internally marking variables that
  represent action choice and relationships between actions), it is
  still hard to write search algorithms as any planning-specific
  insight has to be translated into the CSP representation of
  variables and constraints \comment{For example?}

Constraint-based Planning and specifically Constraint-Based Attribute
and Interval planning (CAIP) \cite{mus94,frank2003} is intended to
close that gap. On one hand, it takes advantage of a CSP
representation; one the other it also uses attributes and intervals to
maintain an explicit representation of the elements relevant to
planning. This makes it easier to write algorithms that search for an
reason abut plans. The main elements of CAIP are:

\begin{enumerate}
	\item Intervals: \comment{needs to be defined}
	\item Attributes:
	\item Domain Constraints an Configuration rules:
\end{enumerate}

\comment{Explain how planning problem instances and their solutions are represented}

### 2012-02-03 related.tex
updates

\begin{wrapfigure}{r}{0.2\textwidth}
 \centering
 \includegraphics[scale=0.3]{figs/planner-arch.pdf}
 \caption{\small A general architectural block diagram for an
   Artificial Intelligence based Planner.}
 \label{fig:planner}
\end{wrapfigure}

@@

The dominant approach for building agent control systems utilize a
$3$-layered architecture (see \cite{gat98} for a well thought out
rationale), notable examples of which include IPEM
\cite{AmbrosIngerson88}, ROGUE \cite{Haigh98}, the LAAS Architecture
\cite{alami:1998p820}, the Remote Agent Experiment \texttt{RAX}
\cite{mus98} and the Autonomous Spacecraft Experiment \texttt{ASE}
\cite{chien99} (see \cite{Knight01} for a survey). Early efforts were
cognizant of if absent with, dealing with time and resources for such
experiments.

>>>>

\begin{figure}[htpb]
 \centering
 \includegraphics[scale=0.3]{figs/planner-arch.pdf}
 \caption{\small A general architectural block diagram for an AI based
   Planner.}
 \label{fig:planner}
\end{figure}

@@

The dominant approach for building agent control systems utilize a
$3$-layered architecture (see \cite{gat98} for a well thought out
rationale), notable examples of which include \texttt{IPEM}
\cite{AmbrosIngerson88}, \texttt{ROGUE} \cite{Haigh98}, the LAAS
Architecture \cite{alami:1998p820}, the Remote Agent Experiment
\texttt{RAX} \cite{mus98} and the Autonomous Spacecraft Experiment
\texttt{ASE} \cite{chien99} (see \cite{Knight01} for a survey). Early
efforts were cognizant of if absent with, dealing with time and
resources for such experiments.
